\documentclass[10pt,a4paper]{article}
\usepackage[margin=1in]{geometry}
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{hyperref}
\usepackage{microtype}

\lstset{
  basicstyle=\ttfamily\small,
  backgroundcolor=\color{black!4},
  frame=single, framesep=4pt, framerule=0pt,
  breaklines=true, columns=flexible,
}

\setlength{\parindent}{0pt}
\setlength{\parskip}{0.5em}

\title{\vspace{-2em}\textbf{Experiment G --- Quadrature Speed/Equivariance Pareto}}
\author{}
\date{}

\begin{document}
\maketitle
\vspace{-3em}

What is the right way to compare DH vs GL? \textbf{Same equivariance error},
then ask how much wall-clock time each method needs to reach it.

\section*{Setup (OC20-scale EquiformerV2, lmax$=$6, mmax$=$2)}

Real \texttt{EquiformerV2Backbone} from fairchem: 12 layers, 128 channels,
24 \texttt{S2Activation} modules. NVIDIA A100-SXM4-40GB, idle.

\begin{center}
\small
\begin{tabular}{llcr}
\toprule
Config       & Method                              & Grid (lat $\times$ lon) & Points \\
\midrule
DH default   & e3nn equiangular                    & 14 $\times$ 5  & 70 \\
DH 2$\times$ & DH with \texttt{resolution=4(L+1)}  & 28 $\times$ 28 & 784 \\
GL min       & custom GL                           & 7 $\times$ 13  & 91 \\
GL match-DH  & custom GL                           & 14 $\times$ 13 & 182 \\
GL 2$\times$ & custom GL                           & 14 $\times$ 28 & 392 \\
\bottomrule
\end{tabular}
\end{center}

\section*{Two measurements at the same (lmax, mmax)}

\textbf{Equivariance error} of one S2-Act kernel (random coefficients,
5 rotations $\times$ 10 inputs); SiLU activation. \emph{Source:}
\texttt{verify\_gl\_grid.py}.

\textbf{Wall-clock} for one full forward (QM9, batch$=$8), 90 measurements
per config (3 runs $\times$ 30 iter), \texttt{cuda.sync()} around each, fresh
model per config. \emph{Source:} \texttt{verify\_speedup\_fairchem\_default.json}.

\section*{Combined results}

\begin{center}
\small
\begin{tabular}{lcrl}
\toprule
Config & Pts & Equiv err & Fwd (ms, 95\% CI) \\
\midrule
DH default                   & 70  & $4.43\!\times\!10^{-1}$ & $111.83 \pm 0.79$ \\
\textbf{DH 2$\times$}        & 784 & $\mathbf{3.27\!\times\!10^{-1}}$ & $\mathbf{162.73 \pm 1.23}$ ($+45.5\%$) \\
GL min                       & 91  & $4.29\!\times\!10^{-1}$ & $\sim$111 (matches DH default) \\
\textbf{GL match-DH}         & 182 & $\mathbf{3.28\!\times\!10^{-1}}$ & $\mathbf{112.23 \pm 0.75}$ ($+0.4\%$ n.s.) \\
\textbf{GL 2$\times$}        & 392 & $\mathbf{3.28\!\times\!10^{-1}}$ & $\mathbf{111.34 \pm 0.76}$ ($-0.4\%$ n.s.) \\
\bottomrule
\end{tabular}
\end{center}

\section*{Matched-equivariance comparison}

For target equiv err $= 3.28 \times 10^{-1}$ (the saturation floor at this mmax cropping):

\begin{itemize}\setlength\itemsep{0pt}
\item Cheapest DH: \textbf{DH 2$\times$ at 162.73 ms}
  (DH default at 70 pts only reaches $4.43\!\times\!10^{-1}$, not enough)
\item Cheapest GL: \textbf{GL match-DH at 112.23 ms}
\item \textbf{Savings: 50.5 ms = $(162.73 - 112.23)/162.73 = 31.0\%$ time reduction}
\end{itemize}

\textbf{So: at matched equivariance accuracy, GL saves $\sim$31\% wall-clock per
forward.}

Equivalently: GL achieves DH 2$\times$'s equivariance at the same wall-clock
as DH default; the $+50.5$ ms premium goes away.

\section*{Caveats}

\begin{enumerate}\setlength\itemsep{0pt}
\item Result is for this specific (lmax$=$6, mmax$=$2) config and this GPU.
  Different model depths or hardware will scale it.
\item Equivariance is measured on \emph{one} S2-Act kernel with random
  inputs, not on a trained model's intermediate features. Trained-model
  features are smoother (absolute equiv numbers will be much lower); the
  \emph{ranking} between quadratures should be preserved.
\item The kernel-level equivariance gap between GL match-DH ($3.28\!\times\!10^{-1}$)
  and DH 2$\times$ ($3.27\!\times\!10^{-1}$) is within rounding --- they really
  are the same accuracy.
\item The mmax$=$2 cropping caps achievable equivariance at $\sim$0.328
  regardless of grid; lower error needs mmax$=$lmax (architectural change).
  At lmax$=$mmax both methods can reach $\sim$4$\times$10$^{-3}$ floor; there
  GL needs roughly half the latitude points of DH (see
  \texttt{verify\_gl\_grid.py}).
\item We did not measure end-to-end equivariance on a fully trained
  OC20-scale model. The matched-accuracy speedup claim rests on the
  kernel-level Pareto.
\end{enumerate}

\section*{Earlier vs revised framing}

The previous report claimed ``DH 2$\times$ costs $+45.5\%$ over DH default'',
which was correct but compared different equivariance levels. The fairer claim:

\begin{quote}
Switching DH 2$\times$ $\rightarrow$ GL match-DH saves \textbf{31\% wall-clock
per forward} while preserving the same equivariance error.
Switching DH default $\rightarrow$ GL match-DH is \textbf{free} (same wall-clock)
and substantially improves equivariance ($4.43\!\times\!10^{-1}
\rightarrow 3.28\!\times\!10^{-1}$).
\end{quote}

\section*{Reproduce}

\begin{lstlisting}
module load pytorch/2.6.0-1
# Equivariance for all configs at lmax=6, mmax=2
python3 experiments/verify_gl_grid.py
# Wall-clock: rigorous 3-run verification
PYTORCH_CUDA_ALLOC_CONF=expandable_segments:True \
  python3 experiments/expG_verify_speedup.py \
    --config fairchem_default --batch_size 8 \
    --n_repeats 3 --n_iter 30 --n_warmup 20
\end{lstlisting}

\end{document}
